Operational constraints and closing civic space: International NGO restrictions and China’s Overseas NGO law

### 2023-12-07 manuscript/manuscript.qmd
Shrink/edit case study section

As explained previously, we expect that the contentiousness of an NGO's issue area should influence its geographic reach and overall flexibility. No INGOs engaged in contentious issues like human rights advocacy have successfully registered in China, but there is some slight variation in the level of contention even among organizations working on less contentious issues like humanitarian assistance, health, and the environment. INGOs that survived the registration thresholds are those already "censored" and deemed to be safe and non-contentious. This can help explain why some seemingly more contentious issue areas such as environmental INGOs enjoy relatively more operational space. Chinese authorities can see the work of environmental INGOs as providing benefits such as political stability and "selectively encourage low risk groups and selectively channel higher risk groups into permitted areas" [@Plantan:2022, 504].

Education is particularly notable as the least flexible and most geographically restricted subsector. While education is typically considered a low-contention issue, in China fears of Western influence through educational institutions were an original impetus for the ONGO law. In 2006, an article in a CCP-owned newspaper warned that the growing presence of international NGOs would "undermine national security," "destroy political stability," "foster corruption," "propagate foreign practices," and "spy on and gather information on China's military, political, and economic information" [@Yin:2009, 534]. The idea that INGOs would "propagate foreign practices" had especially powerful salience, since increased exposure to the West through international NGOs had the potential to "lead to adoption of Western ideas of liberty, further endangering government control over the populace" [@HsiaWhite:2002, 337]. Educational NGOs were one particular avenue for the West to exert influence in the country---the directors and presidents of dozens of China's largest domestic NGOs have received doctoral degrees from and hold visiting appointments at Cornell, Duke, UNC-Chapel Hill, Harvard, Yale, and other prominent US universities [@Wang:2012, 108--9], and their training and connections abroad have influenced their programming and advocacy at home. In 2006, one Chinese NGO leader candidly explained that "foreign influence is definitely great. There is conceptual influence. Foreign NGOs' working methods also affect Chinese NGOs" [@Wang:2012, 110]. Education is therefore one of the more contentious issues allowed under the ONGO law, and thus more closely regulated and restricted.

@@

We also conducted the qualitative study of two selected cases of INGOs in China to further understand the patterns revealed in the quantitative modeling and to examine a more comprehensive picture of how the operations of INGOs have been impacted by the ONGO Law. As explained, the registration documentation data enables us to analyze how the operational space is associated with issue areas and other factors for those INGOs that could successfully register. However, for INGOs that could not get registered but try to maintain their legal presence in China, a typical choice would be constantly filing temporary activities [@Corsetti:2019]. So, we purposefully selected two environmental INGOs with different registration status to conduct a comparative analysis. One of them, the Nature Conservancy (“TNC”), successfully registered a RO under the ONGO Law, while the other INGO, Greenpeace has been filing temporary activities one after another since 2017 to continue carrying out programs in China. We use secondary data including public interviews (previously conducted by one of the authors) and their websites, in addition to official ONGO Law records of ROs and temporary activities to conduct the case studies.

TNC and Greenpeace are both well-known INGOs working in the environmental issue area. Their headquarters are based in western countries and have offices around the world. They started to have long-standing offices in Beijing around the turn of the century, with TNC in 1998 and Greenpeace in 2001. Despite these similarities, the two INGOs had quite different experiences to sustain their legal presence in China after the ONGO Law took effect. According to records by the MPS, TNC’s RO was registered on May 19, 2017, with the National Forestry and Grassland Administration as their PSU. The RO in Beijing is eligible to operate in 27 provinces. Comparatively, Greenpeace has not registered any RO as of December 1, 2023. Instead, they have filed 67 temporary activities, almost going through the filing process 10 times each year.

There are various drawbacks to having to constantly file temporary activities. First would be high administrative costs. Since the filing of temporary activities needs to be submitted by INGOs’ Chinese partners, it would mean discussing and securing the partnerships 67 times, mostly with different organizations. Moreover, this model also imposes high uncertainty to the operation of INGOs. There is no guarantee whether the next Chinese partner would agree to collaborate and file the documentation for them. And it is impossible to run longer-term projects, given all temporary activities cannot last longer than a year. Finally, temporary activities would be programs Chinese partners feel comfortable to collaborate in, i.e., safe activities with little political and administrative risks. So there is less flexibility in the kinds of work an INGO can do with temporary activities.

What factors could have contributed to the contrasting experiences of TNC and Greenpeace under the ONGO Law? After analyzing the data, the working approach and relationship with the government emerged as pivotal differences between the two INGOs, which fits existing discussions on the differential treatment of INGOs based on their relationship with the authority [@Plantan:2022]. In a public interview with TNC’s RO in China about their registration experience, the chief representative stated TNC’s working approach in China has been “science-based”, using “non-confrontational, collaborative approach in working closely with governments, the business community and others” [@ChinaPhilanthropyReserchInstitute:2018a]. Over the years, they have provided technical support for the central government and local governments like the Yunnan and Zhejiang Provinces to build national parks and nature reserve areas and helped drafted such documents as the China Biodiversity Conservation Strategy and Action Plan (2011-2030). Using the non-confrontational working model and establishing long partnerships with the government enabled TNC to secure the consent of a national governmental agency in corresponding issue area to be their PSU, and thus successfully got registered and authorized to operate in a larger number of provinces. In contrast, Greenpeace states in their mission that they work as “independent campaigning organizations that use peaceful protest and creative confrontation to expose global environmental problems”. Their description of the work in China also uses such key words as “brings change”, “non-violent direct action”, and “impartiality and independence”. One project Greenpeace shares on their website was the exposure of the damages the company Sinar Mas Group's did in southwest Yunnan back in 2004. They conducted three rounds of  independent investigations and later revealed that the multinational paper industry company used reforestation as a cover while actually logging Yunnan's natural forests. Although this confrontation was not with the government, it showcases Greenpeace’s investigative, watchdog like work stance. To the authority, it would mean less predictable and manageable.

@@

@tbl-case-study summarizes the comparison of the cases of TNC and Greenpeace. The case study shows that in addition to the correlation of issue area and operational space found in the quantitative model, the dimension of relationship with the government also exerts influence on the contentious level and restrictions INGOs can face, revealing the potential patterns of differential treatment INGOs can receive in the *de facto* implementation of the law.

>>>>

As explained previously, we expect that the contentiousness of an NGO's issue area should influence its geographic reach and overall flexibility. No INGOs engaged in contentious issues like human rights advocacy have successfully registered in China, but there is some slight variation in the level of contention even among organizations working on less contentious issues like humanitarian assistance, health, and the environment. Education is particularly notable as the least flexible and most geographically restricted subsector. While education is typically considered a low-contention issue, in China fears of Western influence through educational institutions were an original impetus for the ONGO law. In 2006, an article in a CCP-owned newspaper warned that the growing presence of international NGOs would "undermine national security," "destroy political stability," "foster corruption," "propagate foreign practices," and "spy on and gather information on China's military, political, and economic information" [@Yin:2009, 534]. The idea that INGOs would "propagate foreign practices" had especially powerful salience, since increased exposure to the West through international NGOs had the potential to "lead to adoption of Western ideas of liberty, further endangering government control over the populace" [@HsiaWhite:2002, 337]. Educational NGOs were one particular avenue for the West to exert influence in the country---the directors and presidents of dozens of China's largest domestic NGOs have received doctoral degrees from and hold visiting appointments at Cornell, Duke, UNC-Chapel Hill, Harvard, Yale, and other prominent US universities [@Wang:2012, 108--9], and their training and connections abroad have influenced their programming and advocacy at home. In 2006, one Chinese NGO leader candidly explained that "foreign influence is definitely great. There is conceptual influence. Foreign NGOs' working methods also affect Chinese NGOs" [@Wang:2012, 110]. Education is therefore one of the more contentious issues allowed under the ONGO law, and thus more closely regulated and restricted.

@@

The results from our model provide useful insight into the operational space of INGOs that have been able to formally register. These organizations have been pre-screened and deemed to be safe and non-contentious. This can help explain why some seemingly more contentious issue areas such as environmental INGOs enjoy relatively more operational space (see @fig-issue-preds). Chinese authorities can benefit politically from the of environmental INGOs and "selectively encourage low risk groups and selectively channel higher risk groups into permitted areas" [@Plantan:2022, 504]. Using our data, however, we are unable to make comparisons with already-vetted INGOs and other potentially more contentious organizations, which limits our analysis of the determinants of operational space. To address this, we compare the cases of two environmental INGOs: (1) The Nature Conservancy (TNC), which successfully registered as RO under the ONGO Law, and (2) Greenpeace, which has filed dozens of sequential temporary activities.

TNC and Greenpeace are both well-known environmental INGOs. Both are based in Western countries and both have offices around the world. They each began working in China around the turn of the century, with TNC in establishing an office in Beijing in 1998 and Greenpeace in 2001. Despite these similarities, the two INGOs have had quite different experiences to sustain their legal presence in China after the ONGO Law took effect. TNC obtained RO status in May 2017, partnering with the National Forestry and Grassland Administration as their PSU, and receiving authorization to work in 27 provinces. In contrast, Greenpeace has not received RO status and has instead filed 67 temporary activities, filing new documentation nearly ten times each year.

There are several drawbacks to having to constantly file temporary activities [@Corsetti:2019], resulting in far more limited INGO operational space than under RO regulations (recall @fig-ro-ta-pathways). Securing a new partnership with a Chinese organization for each activity and co-filing the requisite paperwork imposes substantial administrative costs. This approach also creates high uncertainty, since there is no guarantee that a current partner organization would agree to repeat their supervisory role for the next project. Moreover, INGO programming has an inherent short-term focus, given that temporary activities cannot last longer than a year. Finally, temporary activities need to focus on safe activities with few political risks in order to attract potential Chinese partners.

One key difference in these two organizations' differing registration strategies is their relationship with the government and alignment with government preferences. TNC has stated that its work "science-based" and that the organization uses a "non-confrontational, collaborative approach in working closely with governments, the business community and others. [@ChinaPhilanthropyReserchInstitute:2018a] Over its past two decades of work in China, TNC has provided technical support for the central government and local governments in Yunnan and Zhejiang Provinces, consulting on policies that have led to the establishment of national parks and nature reserves, and helping draft environmental policy proposals like the China National Biodiversity Conservation Strategy and Action Plan (2011–2030). This deliberately non-confrontational model, coupled with long-established partnerships with the government enabled TNC to secure the consent of a environment-focused national governmental agency to act as their PSU, which in turn allowed TNC to have a wider geographic reach throughout the country. Moreover, TNC's work with local and national governments has generally been aligned with government preferences and directly benefits the regime.

In contrast, Greenpeace engages in direct advocacy, "us[ing] peaceful protest and creative confrontation to expose global environmental problems" [@todo]. The organization describes its work in China with phrases like "non-violent direct action," "impartiality and independence," and "brings change", indicating its dedication to maintaining organizational independence as it acts as an environmental watchdog [@todo]. In 2004, for instance, Greenpeace conducted three rounds of independent investigations into damages caused by the Sinar Mas Group, revealing how the Asian multinational corporation used promises of reforestation work as a cover while logging natural forests in Yunnan [@todo]. Greenpeace's ongoing independent advocacy work is often at odds with regime preferences, and from the government's perspective, its Chinese programming is unpredictable and less directly manageable. As a result, few potential PSUs have shown interest in sponsoring any potential registration attempt, and Greenpeace has chosen to not try to permanently register and instead work through repeated temporary activities.

@tbl-case-study summarizes the cases of TNC and Greenpeace. In addition to confirming the correlation between issue area and operational space found in the statistical model, the cases highlight how government preferences exert influence on both the degree of contentious programming INGOs are allowed to undertake and the severity of legal restrictions INGOs can face.

@@
